Hydraulically-vulnerable trees survive on deep-water access in a tropical forest community

### 2020-10-24 manuscript.rmd
1. Explained why ERD is non-coninuous in the community-wide ERD chart

To test whether ERD plays a role in alleviating mortality risk, ee calculated mortality rates for large trees ($\ge$ 10 cm dbh) of species with ERD estimates (all canopy species with maximum height $\ge$30 m) and average abundance of at least 20 trees (one species with an average abundance of 9 trees was eliminated; $n$ = `r length(erd.mrate.sp)`). For each of the seven census intervals $t$ in the 35-yr record (1981-82, 1985, 1990, 2000, 2005, 2010, 2015, 2015), $\textrm{Mort}_{t}$ (% yr^-1^) for species $s$ for two successive censuses, $c_{1}$ and $c_{2}$, was calculated as $\textrm{Mort}_{s,t} = \frac{D_{s,c_{1}}}{N_{s,c_{2}}}\times\frac{100}{\textrm{d}}$, where $N_{s,c_{1}}$ and $D_{s,c_{2}}$ are the total number of trees of a species $s$ present in $c_{1}$ and dead in $c_{2}$, respectively, in the 50 ha plot and $d$ is the duration based on mean dates of $c_{1}$ and $c_{2}$.

@@

\
Figure \@ref(fig:sp-erd): Modelled Effective Rooting Depth (ERD; mean $\pm$ 1SE in m) for `r  length(erd.sp)` large, canopy species of Barro Colorado Island.

>>>>

To test whether ERD plays a role in alleviating mortality risk, we calculated mortality rates for large trees ($\ge$ 10 cm dbh) of species with ERD estimates (all canopy species with maximum height $\ge$30 m) and average abundance of at least 20 trees (one species with an average abundance of 9 trees was eliminated; $n$ = `r length(erd.mrate.sp)`). For each of the seven census intervals $t$ in the 35-yr record (1981-82, 1985, 1990, 2000, 2005, 2010, 2015, 2015), $\textrm{Mort}_{t}$ (% yr^-1^) for species $s$ for two successive censuses, $c_{1}$ and $c_{2}$, was calculated as $\textrm{Mort}_{s,t} = \frac{D_{s,c_{1}}}{N_{s,c_{2}}}\times\frac{100}{\textrm{d}}$, where $N_{s,c_{1}}$ and $D_{s,c_{2}}$ are the total number of trees of a species $s$ present in $c_{1}$ and dead in $c_{2}$, respectively, in the 50 ha plot and $d$ is the duration based on mean dates of $c_{1}$ and $c_{2}$.

@@

\
Figure \@ref(fig:sp-erd): Modelled Effective Rooting Depth (ERD; mean $\pm$ 1SE in m) for `r  length(erd.sp)` large, canopy species of Barro Colorado Island. Each species' ERD is a median of the species' ERDs (depths) obtained from all hydrological realisations, in which soil layers are vertically resolved into exponentially increasing depths. Hence the non-continuous nature of the ERD axis.
